\section{Pillar I: Training Set Generation}
\label{sec:mapping}

A main idea of the proposed method is to shift computational effort to
an offline phase, before the MAV is used on a repetitive task. This
shift is accomplished by using a light-weight machine learning
approach instead of a computationally more complex algorithm, such as
homography finding based on salient image keypoints.

Supervised machine learning needs a training set to find a mapping
from features to target values. In a computer vision-task, the
features represent image characteristics, such as average pixel values
or the amount of oriented gradients.

The proposed method is based on
histograms of textons. To create the training set, a convenient way is
to align image features---that is, the texton histograms---with
high-precision position estimates from a motion tracking system. 
These
systems can track rigid bodies at a high frequency within an error of
few millimeters. This allows to create high-quality training data
sets, where even low quality images can be mapped to the corresponding
$x,y$-position. 

This is achieved by saving the texton histogram on the MAV’s hard disk. The corresponding position from the motion tracking system is broadcast to the UAV via the ground station. The ground station receives the positions from the motion tracking system. 
The major disadvantages of the approach are that
motion tracking systems are usually expensive and time-consuming to
move to different environments.

As an alternative, we sought a low-budget and more flexible
solution. While the homography-based approach (Section xxx) shows high matching quality
for non-blurry images, it suffers from lacking robustness to noise
and the high processing time requirements. To exploit the advantages,
and straighten out the disadvantages, we used the approach in a preprocessing
step to obtain labeled training data of a known environment. This
allows to shift computational effort from the flight phase to a
pre-flight phase---paving the way for autonomous flights of MAVs with
limited processing power. It is based on two associated cornerstones
that are performed in an offline phase: the creation of an
\emph{orthomap} and finding a homography. The required image dataset
for both cornerstones can be obtained by using in-flight pictures of a
manual flight or by taking pictures with a hand-held camera.

For creating the orthomap, the images from the dataset have to be stichted
together to get a hyperspatial image of the scene. By orthorectifying
these pictures, different camera angles can be straightened out,
yielding seamless stitches. This results in a map of the
environment. The stitched image has a higher resolution than the
single images and contains a greater range of detail. The stitching
step includes challenges: a subset of the recorded images might be
distorted and each image must be orthorectified, otherwise perspective
transformations could hinder the stitching process. The images can be
orthorectified by estimating the most probable viewing angle based on
the set of all images. Since a downward-looking camera is attached to
the UAV, most images will be roughly aligned with the z-axis, given
slow flight.

For both steps, the software Agisoft
PhotoScan~\cite{agisoft2013agisoft} was used. The stitching process
can be time-consuming and error-prone. In some environments, an image
from a top view point can be taken capturing the entire area with
high-resolution camera, circumventing the need for stitching together
multiple images. Yet another method starts with an existing image and
modifies the environment accordingly---for example by painting the
floor or printing posters--to correspond to the image. In all cases,
we will describe the environment as \emph{map image}. 

Keypoints of the current image and the map image are detected and
described using the SIFT algorithm. This is followed by a matching
process, that identifies corresponding keypoints between both
images. These matches allow for finding a homography between both
images. For determining the $x, y$-position of the current image, the
center of it is projected on the reference image using the homography
matrix. The pixel position of the center in the reference image can be
used to determine the real world position by transforming the pixel
coordinates to real-world coordinates, based on the scale factors
$C_x$ and $C_y$, with $C_x = \frac{width(R)}{width(I)}$ and
$C_y = \frac{height(R)}{height(I)}$, where $W$ is the real-world
representation and $I$ the digital pixel image. This yields a dataset
of images, labeled with $x, y$ coordinates and the number of
matches. This process already introduces noise into the dataset, since
SIFT can have wrong and inaccurate matches.