\subsection{Evaluation Scheme}
\label{sec:evaluationscheme}

The performance of the proposed method depends
on the environment it is used in: obviously, a texture-rich environment without repetiting patterns will be better suited than an uniformly colored environment. Ideally, one would like to know if the proposed method is able to work in a given environment. Therefore, we propose an evaluation scheme for given
maps. This scheme assigns a global fitness value to a given map,
proportional to the expected accuracy it is expected to achieve in the physical
world. The evaluation scheme allows to inspect the given map and to detect
regions that are responsible for the overall fitness value. The evaluation of a map is
difficult, since the obtained histograms during a real flight depend
on many factors: motion blur, distance to the map and rotations
proportional to the map.

In the first step of the map evaluation procedure, $N$ different
patches of a given map are generated using the tool \emph{draug}
(Section~\ref{sec:syntheticdatageneration}). We propose the following global loss function
($L$) for evaluating a given map ($\mathcal{M}$):

\begin{align}
  L(\mathcal{M}) &= \sum_{i = 1}^{N} \sum_{j = 1}^{N} \ell(d_a(h_i, h_j), d_e(h_i, h_j))
\end{align}

\begin{align}
  \ell(x, y) &= x - y\\
  d_a(h_i, h_j) &= \text{CS}(h_i, h_j)\\
  d_e(h_i, h_j) &= f_X(pos_i \mid \mu, \Sigma) = f_X(x_i, y_i \mid \mu, \Sigma)\\
\end{align}

\begin{align}
\mu = pos_j = (x_j, y_j)\\
\Sigma =
  \begin{bmatrix}
    \rho & 0\\
    0 & \rho\\
  \end{bmatrix}
\end{align}

The cosine similarity (CS) is defined as:
\begin{align}
CS(h_i, h_j) = \frac{h_i^Th_j}{||h_i||||h_j||}
\end{align}

The cosine similarity has the convenient property that its values are bounded between $-1$ and $1$. In the present case, since the elements of $h_i$ and $h_j$ are non-negative, it is even bounded between $0$ and $1$. The function $f_x$ describes the non-normalized probability density function of the normal distribution: $f_x = e^{- \frac{(x - \mu)^2}{2 \sigma ^ 2}}$. This function is also bounded between $0$ and $1$, which makes the functions $f_X$ and $CS$ easily comparable.

The idea behind the global loss function $L$ is that histograms in closeby areas
should be similar and the similarity should decrease the further away
two positions are. This is modeled as a 2-dimensional Gaussian with zero
covariance. The variance is depended on the
desired accuracy ($\rho$): the lower the variance, the more punctuated
a certain location is but also the higher the risk that a totally
wrong measurement occurs. The following visualization are based on
color histograms (and not texton histograms) for easier visual
analysis.